An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-02-16 EconGrowth.Rnw
Edited the CES and LINEX equations.

\begin{equation} \label{eq:Single_Factor_Generic}
  y = a^{\lambda (t-t_0)}f^{m}
\end{equation}

@@

\begin{equation} \label{eq:CD_No_Energy}
  y = a^{\lambda (t-t_0)}k^{\alpha}l^{\beta}.
\end{equation}

@@

\begin{equation} \label{eq:CD_With_Energy}
  y = a^{\lambda (t-t_0)}k^{\alpha}l^{\beta}e^{\gamma},
\end{equation}

@@

<<CES functions, eval=FALSE>>=
####################
# Returns an nls ces model for the country and energyType specified.
# energyType should be one of Q", "X", or "U".
##
cesModel <- function(countryAbbrev, energyType){
  # Load the data that we need.
  dataTable <- loadData(countryAbbrev)
  # We need to do the CES fit with the desired energyType.
  # To achieve the correct fit, we'll change the name of the desired column
  # to "iEToFit" and use "iEToFit" in the nls function.
  dataTable <- replaceColName(dataTable, energyType, "iEToFit")
print(dataTable)
  # Establish guess values for phi beta, zeta, lambda_L and lambda_E.
  sigmaGuess <- 0.663 # 0.663 for Sweden
  zetaGuess <- 0.147 # 0.147 for Sweden
  betaGuess <- 0.7 # 0.7 for Sweden
  lambda_LGuess <- 0.011 # 0.011 for Sweden
  lambda_EGuess <- 0.018 # 0.018 for Sweden
  # Runs a non-linear least squares fit to the data with constraints
  modelCES <- nls(iGDP ~ ((1-zeta)^(1.0/sigma) * ((exp(lambda_L*iYear))^beta 
                                                  * iCapStk^(1-beta) 
                                                  * iLabor^beta)^((sigma-1.0)/sigma) 
                          + zeta^(1.0/sigma)*(exp(lambda_E*iYear) 
                                                  * iEToFit)^((sigma-1.0)/sigma))^(sigma/(sigma-1.0)), 
                  data=dataTable,
                  start=list(sigma=sigmaGuess, zeta=zetaGuess, beta=betaGuess, 
                             lambda_L=lambda_LGuess, lambda_E=lambda_EGuess),
                  control=nls.control(maxiter=10000, tol=1e-06, minFactor=1/1024, 
                                      printEval=FALSE, warnOnly=TRUE),
                  algorithm = "port",
                  lower = list(sigma=-Inf, zeta=0, beta=0.0, lambda_L=0.0, lambda_E=0.0),
                  upper = list(sigma= Inf, zeta=1, beta=1.0, lambda_L=0.1, lambda_E=0.1),
                  )
  return(modelCES)
}
model <- cesModel("US", "Q")
summary <- summary(model)
print(summary)

@@

@@

\begin{equation} \label{eq:LINEX_model}
  y = e \: \mathrm{exp} \left[a_0 \left(2 - \frac{l + e}{k} \right) + a_0 c_t \left(\frac{l}{e} - 1 \right) \right],
\end{equation}

\noindent where $e$ is any of $q$, $x$, or $u$, and $a_0$ and $c_t$ are fitting parameters.

@@

<<LINEXq_Params_Graph, fig.width=linexParameterGraphWidth, fig.height=linexParameterGraphHeight, fig.pos="H", fig.cap="LINEX (with $q$) model parameters. Vertical bars indicate 95\\% confidence intervals.", eval=TRUE>>=
createLINEXParamsGraph("Q")
@

@@

<<LINEXx_Params_Graph, fig.width=linexParameterGraphWidth, fig.height=linexParameterGraphHeight, fig.pos="H", fig.cap="LINEX (with $x$) model parameters. Vertical bars indicate 95\\% confidence intervals.", eval=TRUE>>=
createLINEXParamsGraph("X")
@

@@

<<LINEXu_Params_Graph, fig.width=linexParameterGraphWidth, fig.height=linexParameterGraphHeight, fig.pos="H", fig.cap="LINEX (with $u$) model parameters. Vertical bars indicate 95\\% confidence intervals.", eval=TRUE>>=
createLINEXParamsGraph("U")
@

>>>>

\begin{equation} \label{eq:Single_Factor_Generic}
  y = \mathrm{e}^{\lambda (t-t_0)}f^{m}
\end{equation}

@@

\begin{equation} \label{eq:CD_No_Energy}
  y = \mathrm{e}^{\lambda (t-t_0)}k^{\alpha}l^{\beta}.
\end{equation}

@@

\begin{equation} \label{eq:CD_With_Energy}
  y = \mathrm{e}^{\lambda (t-t_0)}k^{\alpha}l^{\beta}e^{\gamma},
\end{equation}

@@

<<CES functions, eval=TRUE>>=
####################
# Returns an nls ces model for the country and energyType specified.
# energyType should be one of Q", "X", or "U".
##
cesModel <- function(countryAbbrev, energyType){
  # Load the data that we need.
  dataTable <- loadData(countryAbbrev)
  # We need to do the CES fit with the desired energyType.
  # To achieve the correct fit, we'll change the name of the desired column
  # to "iEToFit" and use "iEToFit" in the nls function.
  dataTable <- replaceColName(dataTable, energyType, "iEToFit")
  # Establish guess values for phi beta, zeta, lambda_L and lambda_E.
  sigmaGuess <- 0.663 # 0.663 for Sweden
  # gamma_EGuess <- 0.147 # 0.147 for Sweden
  gamma_EGuess <- 0.01
  betaGuess <- 0.7 # 0.7 for Sweden
  lambda_LGuess <- 0.011 # 0.011 for Sweden
  lambda_EGuess <- 0.018 # 0.018 for Sweden
  # Runs a non-linear least squares fit to the data with constraints
  modelCES <- nls(iGDP ~ ((1-gamma_E)^(1.0/sigma) * ((exp(lambda_L*iYear))^beta 
                                                  * iCapStk^(1-beta) 
                                                  * iLabor^beta)^((sigma-1.0)/sigma) 
                          + gamma_E^(1.0/sigma)*(exp(lambda_E*iYear) 
                                                  * iEToFit)^((sigma-1.0)/sigma))^(sigma/(sigma-1.0)), 
                  data=dataTable,
                  start=list(sigma=sigmaGuess, gamma_E=gamma_EGuess, beta=betaGuess, 
                             lambda_L=lambda_LGuess, lambda_E=lambda_EGuess),
                  control=nls.control(maxiter=10000, tol=1e-06, minFactor=1/1024, 
                                      printEval=FALSE, warnOnly=TRUE),
                  algorithm = "port",
                  lower = list(sigma=-Inf, gamma_E=0.0, beta=-Inf, lambda_L=-Inf, lambda_E=-Inf),
                  upper = list(sigma= Inf, gamma_E=1.0, beta= Inf, lambda_L= Inf, lambda_E= Inf),
                  )
  return(modelCES)
}
print("US Q")
print(summary(cesModel("US", "Q")))

print("")
print("UK Q")
print(summary(cesModel("UK", "Q")))

print("")
print("JP Q")
print(summary(cesModel("JP", "Q")))

print("")
print("CN Q")
print(summary(cesModel("CN", "Q")))

print("")
print("ZA Q")
print(summary(cesModel("ZA", "Q")))

print("")
print("SA Q")
print(summary(cesModel("SA", "Q")))

print("")
print("IR Q")
print(summary(cesModel("IR", "Q")))

print("")
print("TZ Q")
print(summary(cesModel("TZ", "Q")))

print("")
print("ZM Q")
print(summary(cesModel("ZM", "Q")))

@@

The CES model is of the form

\begin{equation} \label{eq:CES_model}
  y = \left\{ (1-\gamma_E)^{1/\sigma} \left[ k^{\alpha} \left( \mathrm{e}^{\lambda_L(t-t_0)} l \right)^{\beta} \right]^{\frac{\sigma-1}{\sigma}}   
      + (\gamma_E)^{1/\sigma} \left[ \mathrm{e}^{\lambda_E(t-t_0)} e \right]^{\frac{\sigma - 1}{\sigma}}   \right\}
      ^{\frac{\sigma}{\sigma - 1}},
\end{equation}

\noindent where $e$ is any of $q$, $x$, or $u$; $\gamma_E$, $\lambda_L$, $\lambda_E$, $\beta$, and $\sigma$ are fitting parameters; and $\mathrm{e}$ is the base of the natural logarithm. Ths sum $\alpha + \beta$ is constrained to unity.

@@

\begin{equation} \label{eq:LINEX_model}
  y = e \: \mathrm{e}^{\left[a_0 \left(2 - \frac{l + e}{k} \right) + a_0 c_t \left(\frac{l}{e} - 1 \right)\right]} ,
\end{equation}

\noindent where $e$ is any of $q$, $x$, or $u$; $a_0$ and $c_t$ are fitting parameters; and $\mathrm{e}$ is the base of the natural logarithm.

@@

<<LINEXq_Params_Graph, fig.width=linexParameterGraphWidth, fig.height=linexParameterGraphHeight, fig.pos="H", fig.cap="LINEX (with $q$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createLINEXParamsGraph("Q")
@

@@

<<LINEXx_Params_Graph, fig.width=linexParameterGraphWidth, fig.height=linexParameterGraphHeight, fig.pos="H", fig.cap="LINEX (with $x$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createLINEXParamsGraph("X")
@

@@

<<LINEXu_Params_Graph, fig.width=linexParameterGraphWidth, fig.height=linexParameterGraphHeight, fig.pos="H", fig.cap="LINEX (with $u$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createLINEXParamsGraph("U")
@

\subsection{LINEX comparisons}
